---
title: "Final project code"
author: "Pei Gong"
date: "December 2, 2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stringr)
library(rpart)
library(purrr)
library(igraph)
```

#Read in the data 
```{r}
orders <- read.csv("orders.csv")
products_prior <- read.csv("order_products__prior.csv")
order_products_train <- read.csv("order_products__train.csv")
products <- read.csv("products.csv")
aisles <- read.csv("aisles.csv")
departments <- read.csv("departments.csv") 
```

#Exploratory data analysis 

## Most ordered product by departments
```{r}
#Here, we join the orders by time with the products ordered for prior and train set orders
prior <- products_prior %>% left_join(orders, by = "order_id")

train <- order_products_train %>% left_join(orders, by = "order_id")
#Joining the prior and training sets together
orders_full <- rbind(train, prior) 

orders_full<-subset(orders_full,eval="test")

products <- products %>% 
    mutate(organic=ifelse(str_detect(str_to_lower(products$product_name),'organic'),1,0), organic= as.factor(organic)) 

orders_1<-orders_full %>% left_join(products, "product_id") 
 
by_user_summary<-orders_1 %>% group_by(user_id,product_name) 

by_department_summary <- orders_1 %>% group_by(department_id, product_name) %>%
  summarise(count = n())
#Finding the most ordered product per hour and joining with product names
by_department_max <- by_department_summary %>% group_by(department_id) %>% filter(count == max(count)) %>%  left_join(departments, "department_id")
department1<-by_department_max[,c(2,4)]
```

## Most ordered product by aisle

```{r}
#Counting how many of each product is ordered each day (notice use 
# of orders_full made above)
by_aisle_summary <- orders_1 %>% group_by(aisle_id, product_name) %>%
  summarise(count = n())
#Finding the most ordered product per hour and joining with product names
by_aisle_max <- by_aisle_summary %>% group_by(aisle_id) %>% filter(count == max(count)) %>% left_join(aisles, "aisle_id")
by_aisle1<-by_aisle_max[,c(2,4)]
```

##Most ordered product by hour (ignoring day of week)
```{r}
#Here, we join the orders by time with the products ordered for prior and train set orders
prior <- products_prior %>% left_join(orders, by = "order_id")
train <- order_products_train %>% left_join(orders, by = "order_id")
#Joining the prior and training sets together
orders_full <- rbind(train, prior)
#Counting how many of each product is ordered in each hour
by_hour_summary <- orders_full %>% group_by(order_hour_of_day, product_id) %>%
  summarise(count = n())
#Finding the most ordered product per hour and joining with product names
by_hour_max <- by_hour_summary %>% group_by(order_hour_of_day) %>% filter(count == max(count)) %>% left_join(products, "product_id")
#Graphing results
ggplot(data=by_hour_max, aes(x=order_hour_of_day, y=count, fill=product_name)) +
    geom_bar(stat="identity") + xlab("Hour of Day") + ylab("Count") + 
  ggtitle("Most Ordered Product Per Hour") + guides(fill=guide_legend(title="Product"))
```

## Most ordered product by day of week
```{r}
#Counting how many of each product is ordered each day (notice use 
# of orders_full made above)
by_day_summary <- orders_full %>% group_by(order_dow, product_id) %>%
  summarise(count = n())
#Finding the most ordered product per hour and joining with product names
by_day_max <- by_day_summary %>% group_by(order_dow) %>% filter(count == max(count)) %>% left_join(products, "product_id")
#Graphing results
ggplot(data=by_day_max, aes(x=order_dow, y=count, fill=product_name)) +
    geom_bar(stat="identity") + xlab("Day of Week") + ylab("Count") + 
  ggtitle("Most Ordered Product Per Day") + guides(fill=guide_legend(title="Product"))
```

## Most ordered by hour by day
```{r}
#Counting how many of each product is ordered each day by hour (notice use of 
# orders_full made above)
by_day_by_hour_summary <- orders_full %>% 
  group_by(order_dow, order_hour_of_day, product_id) %>% 
  summarise(count = n())
#Finding the most ordered product each hour grouped by day
by_day_by_hour_max <- by_day_by_hour_summary %>% 
  group_by(order_dow, order_hour_of_day) %>% filter(count == max(count)) %>%
  left_join(products, "product_id")
#Didn't graph because it's still all bananas
```

## Predicting banana for every order
```{r}
test <- orders %>% filter(eval_set == "test")
test <- test %>% mutate(products = 24852)
test_to_submit <- test %>% select(order_id, products)
write.csv(test_to_submit, "submission.csv", row.names=FALSE)
```

## Making frequency based visualizations

20 most ordered products period
```{r}
#Tallying product orders
most_ordered <- orders_full %>% group_by(product_id) %>% summarise(count = n()) 
#Arranging by count
most_ordered <- most_ordered %>% arrange(count) 
#Subsetting the top 50 most frequently ordered products
most_ordered <- most_ordered %>% top_n(20)
#Joining with product names
most_ordered <- most_ordered %>% left_join(products, "product_id") 
most_ordered <- most_ordered %>% left_join(departments, "department_id")
#Graphing results
ggplot(data=most_ordered, aes(x=reorder(product_name, -count), y=count, fill = department)) + geom_bar(stat="identity") + coord_flip() + xlab("Product") + ylab("Count") + 
  ggtitle("20 Most Ordered Products") + guides(fill=guide_legend(title = "Department"))+ scale_fill_brewer(palette="YlGn")
```

## Most re-ordered products
```{r}
#Tallying product orders
most_reordered <- orders_full %>% filter(reordered == 1) %>% group_by(product_id) %>% summarise(count = n()) 
#Arranging by count
most_reordered <- most_reordered %>% arrange(count) 
#Subsetting the top 50 most frequently ordered products
most_reordered <- most_reordered %>% top_n(20)
#Joining with product names
most_reordered <- most_reordered %>% left_join(products, "product_id") 
most_reordered <- most_reordered %>% left_join(departments, "department_id")
#Graphing results
ggplot(data=most_reordered, aes(x=reorder(product_name, -count), y=count, fill = department)) + geom_bar(stat="identity") + coord_flip() + xlab("Product") + ylab("Count") + 
  ggtitle("20 Most ReOrdered Products") + guides(fill=guide_legend(title = "Department"))+ scale_fill_brewer(palette="YlGn")
```

## Making a CART tree per user for a small subset of train users
```{r}
#Making subset
set.seed(76)
train_subset <- train %>% sample_n(100)

#Making function to find user id in prior and make tree based on those orders
get_tree <- function(order_id_for_model){
  #Save model formula
  model_formula <- as.formula(as.factor(product_id) ~ order_number + order_dow + order_hour_of_day + days_since_prior_order + reordered)
  
  #Pulling user_id row from train_subset
  train_for_model <- train_subset %>% filter(order_id == order_id_for_model)
  
  #Find user's order in set of prior orders
  data_for_user <- prior %>% filter(user_id == train_for_model$user_id)
  
  #Make tree for user
  model_CART <- rpart(model_formula, data = data_for_user)

  
  #Make predictions
  y_hat <- model_CART %>% predict(newdata=train_for_model, type = "class")
  return(y_hat %>% unlist() %>% as.character() %>% as.numeric())
}

order_id_vector_train_subset <- as_vector(train_subset$order_id)

y_hat_train_subset1 <- rep(0,100)

for(i in 1:100){
  y_hat_train_subset1[i] <- get_tree(order_id_vector_train_subset[i])
}

#Finding raw number correct
train_subset <- bind_cols(train_subset, data.frame(y_hat_train_subset1))
num_correct <- train_subset %>% mutate(correct = ifelse(product_id == y_hat_train_subset1, 1, 0)) %>% summarise(count = sum(correct))

#Question for Office Hours: F1 score calculation in this case?

#False positive is the number wrong of the number in the cart
#True positive is what you get right
```


#model1 
## Predict that an individual will order the item they have ordered most

First, look at what the maximum number of times a user has ordered the same product.

```{r}
orders_full %>% group_by(user_id) %>% count(product_id) %>% summarise(max_orders = max(n)) %>% 
  ggplot() +
  geom_histogram(aes(x = max_orders), fill = "green4") +
  labs(x = "Maximum number of times a user has ordered the same product",
       title = "")
#ggsave("max_reorders.png")
```

Now use predict the next product that a user will order is product that each user has ordered most.

```{r}
max_ordered_product <- orders_full %>% group_by(user_id) %>% 
  count(product_id) 
max_ordered_product <- max_ordered_product %>% filter(n == max(n))

# predict; when a user has multiple products tied for most ordered, break at random
test <- orders %>% filter(eval_set == "test") %>% left_join(max_ordered_product, by = "user_id") 
test <- test %>% group_by(user_id) %>% sample_n(1) %>% ungroup()
  
test <- test %>% select(order_id, product_id) %>% rename(products = product_id)
write_csv(test, "max_previously_ordered.csv")
```

Kaggle F1 score: 0.1632167. 


Now retry but guessing top 2 items.

```{r}
# set up data frame to use to predict 2 or more products
max_ordered_product2 <- orders_full %>% group_by(user_id) %>% 
  count(product_id) %>% top_n(2, wt = n)
max_ordered_product2 <- max_ordered_product2 %>% select(-n)
max_ordered_product_2_ormore <- max_ordered_product2 %>% mutate(row = row_number()) %>% 
  spread(row, product_id)

test2 <- orders %>% filter(eval_set == "test") %>% left_join(max_ordered_product_2_ormore, by = "user_id")
test2 <- test2 %>% select(order_id, '1', '2')
  
test2 <- within(test2, x <- paste(test2$`1`,test2$`2`,sep=' ')) %>% rename(order_id = order_id, products = x) %>% 
  select(order_id, products)
write_csv(test2, "max_previously_ordered_2.csv")
```

Top 3 items.

```{r}
test3 <- orders %>% filter(eval_set == "test") %>% left_join(max_ordered_product_2_ormore, by = "user_id")
test3 <- test3 %>% select(order_id, '1', '2', '3')
test3$`3`[is.na(test3$`3`)] <- ""
  
test3 <- within(test3, x <- paste(test3$`1`,test3$`2`,test3$`3`,sep=' ')) %>% 
  rename(order_id = order_id, products = x) %>% 
  select(order_id, products)


write_csv(test3, "max_previously_ordered_3.csv")
```

Top 4 items.

```{r}
test4 <- orders %>% filter(eval_set == "test") %>% left_join(max_ordered_product_2_ormore, by = "user_id")
test4 <- test4 %>% select(order_id, '1', '2', '3','4')
test4$`3`[is.na(test4$`3`)] <- ""
test4$`3`[is.na(test4$`4`)] <- ""
  
test4 <- within(test4, x <- paste(test4$`1`,test4$`2`,test4$`3`,test4$`4`,sep=' ')) %>% 
  rename(order_id = order_id, products = x) %>% 
  select(order_id, products)


write_csv(test4, "max_previously_ordered_4.csv")

```

# Network

Here we consider that each product is a node. Each connection runs between products that are in the same order. 

```{r}
# create edgelist matrix for just 5 orders
edgelist <- data.frame()

set.seed(11)
orders_full_small <- orders_full %>% 
  filter(order_id %in% sample(orders_full$order_id, size = 5, replace = T)) %>% 
  select(order_id, product_id)
orders_full_small$order_id <- as.factor(orders_full_small$order_id)

# Setting up edgelist matrix as a n x n matrix (n = 62)

all_product_ids <- unique(orders_full_small$product_id)
edgelist_matrix <- matrix(nrow = length(all_product_ids),
                          ncol = length(all_product_ids),
                          dimnames = list(all_product_ids, all_product_ids))

for (x in levels(orders_full_small$order_id)){
  for (i in rownames(edgelist_matrix)){
    for (j in colnames(edgelist_matrix)){
      if (i %in% unique(orders_full_small$product_id[orders_full_small$order_id == x])){
        if (j %in% unique(orders_full_small$product_id[orders_full_small$order_id == x])){
          edgelist_matrix[i,j] <- 1
        }
        else edgelist_matrix[i,j] <- 0
      }
    }
  }
}

# rename rows
rnames <- rownames(edgelist_matrix) %>% 
  as.data.frame() %>% 
  rename(product_id = '.') %>% 
  mutate(product_id = as.integer(product_id)) %>% 
  left_join(products, by = "product_id") %>% 
  select(product_name)
rnames <- rnames$product_name
rnames <- as.character(rnames)

rownames(edgelist_matrix) <- rnames

# rename columns
cnames <- colnames(edgelist_matrix) %>% 
  as.data.frame() %>% 
  rename(product_id = '.') %>% 
  mutate(product_id = as.integer(product_id)) %>% 
  left_join(products, by = "product_id") %>% 
  select(product_name)
cnames <- cnames$product_name
cnames <- as.character(cnames)

colnames(edgelist_matrix) <- cnames
```

What products do we have here?

```{r}
rownames(edgelist_matrix) %>% 
  as.data.frame() %>% 
  rename(product_id = '.') %>% 
  mutate(product_id = as.integer(product_id)) %>% 
  left_join(products, by = "product_id") %>% 
  select(product_name)
```

## Visualize network

```{r}
igraph_adj_network <- graph_from_adjacency_matrix(edgelist_matrix, mode = "undirected", weighted = TRUE,
                                                  add.colnames = 'label')

# function borrowed from Stack Overflow
# https://stackoverflow.com/questions/38999656/increasing-spaces-between-vertices-for-r-igraph
layout.by.attr <- function(graph, wc, cluster.strength=1,layout=layout.auto) {  
  g <- graph.edgelist(get.edgelist(graph)) # create a lightweight copy of graph w/o the attributes.
  E(g)$weight <- 1
  
  attr <- cbind(id=1:vcount(g), val=wc)
  g <- g + vertices(unique(attr[,2])) + igraph::edges(unlist(t(attr)), weight=cluster.strength)
  
  l <- layout(g, weights=E(g)$weight)[1:vcount(graph),]
  return(l)
}

set.seed(9) # set seed because orientation of network changes each time
plot(simplify(igraph_adj_network),
     vertex.size = 5,
     vertex.label = ifelse(degree(igraph_adj_network) > 30, V(igraph_adj_network)$label, NA),
     vertex.label.cex = 1,
     vertex.label.font = 2,
     vertex.label.dist = 2,
     vertex.label.color = "black",
     vertex.color = "green4",
     edge.color = "gray81",
     layout=layout.by.attr(igraph_adj_network, wc=1))
title("Network for 5 Instacart orders",
      cex.main = 2)
```

